Critical assessment of Linear-Response Excitations in Molecular system\\ or \\ On the Localization of the Linear-Response excitations in Open Systems

### 2018-01-24 LRDraft.tex
update

The capability of the computational basis to express the results of the linear response can be discussed on the localization argument previously presented. 
The first step needed to address this problem concerns with the evaluation of the reliability of the $\mu$-basis for $\omega=0$, \emph{i.e.} for a perturbing field that does 
not depend on time. In this case the fluctuation states, as well as the expectation value of a given observable, can be computed performing a ground state calculation in which the perturbing 
field $\op\Phi$ is added to the external potential of $\hnot$\footnote{We are interested in the linear response, so the magnitude of $\op\Phi$ has to be chosen in a way that respects this 
regime. To check this point we can compute the perturbation-induced modification of an observable an verify that its value scales linearly with the magnitude of the field}. 
The goodness of the $\mu$-basis can thus be verified as in the standard ground state calculation, for instance by checking that, given a certain level of completeness of the basis,
the energy of the occupied (perturbed) orbitals is stable with respect to further refinements.

@@

We introduce a set of basis vectors $\{\ket{\psi(\alpha)}\}$ in the subspace of the empty-states, \emph{i.e.} in the subspace orthogonal to the occupied eigenstates of $\hnot$. 
The elements of this set satisfy, by definition, the following orthonormality relation: 
\be\lb{psiAlphaDef1}%{orthonormalityDef1}
\brket{\psi(\alpha)}{\psi_p} = \brket{\psi_p}{\psi(\alpha)} = 0 \;, \qq
\brket{\psi(\alpha)}{\psi(\beta)} = \delta(\alpha-\beta) \;,
\ee
where we have parametrized the basis element through the multi-index $\alpha$, in the sense described in \ref{BasicDef}, since the associated vector can describe both a localized or a 
de-localized state. The $\op Q_0$ projector as well as the completeness relation can be easily written in function of this basis: 
\be\lb{psiAlphaQ0}
\op Q_0 = \sint\dd\alpha\ketbra{\psi(\alpha)}{\psi(\alpha)} \;, \;\; 
\identity = \sum_p\ketbra{\psi_p}{\psi_p} + \sint\dd\alpha\ketbra{\psi(\alpha)}{\psi(\alpha)} \;.
\ee
The dimension $n_\alpha$ of the $\alpha$-basis can be determined by expressing its elements $\{\ket{\psi(\alpha)}\}$ in the same $\mu$-basis used to perform the ground states calculation. 
%both with and without the explicit time-independent perturbing field. 
%Denoting with $n_{occ}$ the number of occupied orbitals  
Since the orthogonality condition with respect to the occupied orbitals imposes $n_{occ}$ constraints we conclude that $n_\alpha=n_\mu-n_{occ}$.

@@

A generic operator can be decomposed in this basis and its formal expansion contains four distinct contribution of the form $\ketbra{\psi_p}{\psi_q}$, 
$\ketbra{\psi(\alpha)}{\psi(\beta)}$, $\ketbra{\psi(\alpha)}{\psi_p}$, $\ketbra{\psi_p}{\psi(\alpha)}$. However an inspection of equations \eqref{rhoPrimeOmegaDef1}
and \eqref{LiouvillianRhopomegaDef3} evidences that the relevant operators in the contest of the linear response are expressed as a commutator with $\dmnot$. 
For this class of operators the only non vanishing contribution comes from the \emph{transition operators}:
\be
\excite{p}{\alpha} \equiv \ketbra{\psi(\alpha)}{\psi_p}\;, \qq
\decay{\alpha}{p} \equiv \ketbra{\psi_p}{\psi(\alpha)} \;.
\ee
The orthonormality conditions \eqref{psiAlphaDef1} imply that:
\be\lb{transitionOpOrthonormalityDef1}
\trace{\excite{p}{\alpha}\excite{q}{\beta}} = \trace{\decay{\alpha}{p}\decay{\beta}{q}} = 0 \;, \qq
\trace{\excite{p}{\alpha}\decay{\beta}{q}} = \delta_{pq}\delta(\alpha-\beta) \;,
\ee
so that the expansion of a suitable operator $\op O$ in the basis of transitions reads:
\be\lb{OperatorTransBasisExpDef1}
\op O = \sum_p\sint \dd \alpha\left(
\excite{p}{\alpha}\trace{\decay{\alpha}{p}\op O} + \decay{\alpha}{p}\trace{\excite{p}{\alpha}\op O}
\right) \;.
\ee
In order to recast this expansion in a simpler expression it turns out to be useful to define the rank-2 row vectors of the transition basis: 
\be
\sbra{p\alpha}{\cdot} \equiv 
\mat{ \excite{p}{\alpha} \;, & \decay{\alpha}{p} }\;,
\ee
and we also introduce a vectorial notation to deal with the coefficients of the expansion of an operator in the transition basis, that is: 
\be\lb{OperatorTransitionCoeffDef1}
\sket{p\alpha}{\op O}  \equiv \mat{ \trace{\decay{\alpha}{p}\op O } \\ 
\trace{\excite{p}{\alpha}\op O}} \; .
\ee
Thanks to this formalism we have that equation \eqref{OperatorTransBasisExpDef1} can be expressed as:
\be\lb{OperatorTransBasisExpDef2}
\op O = \sum_p\sint \dd \alpha \sbra{p\alpha}{\cdot}\sket{p\alpha}{\op O} = 
\dbraket{\cdot}{\op O} \;,
\ee
where the last equality implicitly define the ulterior compact notation of the ``big-shaped'' vectors in which the summed indexes are omitted.  
In the same fashion it will be also useful for the subsequent analysis to introduce a ``conjugate'' notation, in which the basis operators are expressed
as column vector and the coefficients of the expanded operator as row ones, that is: 
\be\lb{TransitionBasisConjugateDef1}
\sket{p\alpha}{\cdot} \equiv 
\mat{\decay{\alpha}{p} \\  \excite{p}{\alpha}} \;, \qq
\sbra{p\alpha}{\op O}  \equiv 
\mat{ \trace{\excite{\alpha}{p}\op O } \;, & \trace{\decay{p}{\alpha}\op O} }
\;,
\ee
where the flip in the role of $\op e$ and $\op d$ in \eqref{TransitionBasisConjugateDef1} follows from the properties of excitation and decay operators
under hermitian-conjugate operation. The expansion of a suitable operator in the transition basis using the conjugate description reads: 
\be\lb{OperatorTransBasisConjugateExpDef1}
\op O = \sum_p\sint \dd \alpha \sbra{p\alpha}{\op O}\sket{p\alpha}{\cdot} = 
\dbraket{\op O}{\cdot} \;.
\ee

@@

The Liouville equation \eqref{LiouvillianRhopomegaDef3} can be consistently written in the transition basis. Indeed, its r.h.s is explicitly written
as a commutator with $\dmnot$, while an inspection of the formal definition of the response density and of the Liouville superoperator evidences that
they both contain a commutator with $\dmnot$. Written in the basis of transitions Liouville equation becomes a linear matrix equation for the rank-2 
vectors of the coefficients \eqref{OperatorTransitionCoeffDef1} of the operators in its left and right side:
\be
\omega\mat{ \trace{\decay{\alpha}{p}\dm'(\omega)} \\ \trace{\excite{p}{\alpha}\dm'(\omega)}} - 
\mat{ \trace{\decay{\alpha}{p}\Liouv\dm'(\omega)} \\ \trace{\excite{p}{\alpha}\Liouv\dm'(\omega)}} =
\mat{ \trace{\decay{\alpha}{p}\commutator{\op\Phi(\omega)}{\dmnot}} \\ \trace{\excite{p}{\alpha}\commutator{\op\Phi(\omega)}{\dmnot}}}
\;.\nn
\ee
The second addend in the l.h.s can be further analyzed by writing the transition basis expansion of $\dm'$ inside the traces. Performing this procedure
we can recast this quantity as the action of a 2-dimensional matrix on the rank-2 vector of the response density. So we have: 
\be\lb{LiouvilleTransBasisDef1}
\sum_q\sint\dd\beta \left[\omega\delta_q^p\delta(\alpha-\beta)- \optr{L}^{p \alpha}_{q\beta} \right]\sket{q\beta}{\dm'(\omega)} =
\sket{p\alpha}{\commutator{\op\Phi(\omega)}{\dmnot}} \;.
\ee
with:
\be\lb{LiouvilleMatrixTransBasisDef1}
\optr{L}^{p \alpha}_{q\beta} \equiv \mat{\trace{\decay{\alpha}{p}\, \Liouv\, \excite{q}{\beta}} &
\trace{\decay{\alpha}{p}\, \Liouv\, \decay{\beta}{q}} \\
\trace{\excite{p}{\alpha}\, \Liouv\, \excite{q}{\beta}} &
\trace{\excite{p}{\alpha}\, \Liouv\, \decay{\beta}{q}}} \;.
\ee

@@

This formulation of the Liouville equation is the starting point for the derivation of the spectral decomposition of the response density operator
in terms of the eigenvectors if the Liouville operator. 
The first step in the achievement of this result is provided by the construction of the eigenvectors of the matrix \eqref{LiouvilleMatrixTransBasisDef1}.
Following the notation previously defined we introduce the rank-2 vectors $\sket{p\alpha}{E}$ and $\sbra{p\alpha}{\tilde E}$ that represent, by definition,
the right and left eigenvectors $\optr{L}$, respectively, so that:
\be\lb{LiouvMatrixEigenvectorDef1}
\sum_q \sint \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q \beta}{E} = E \sket{p\alpha}{E} \;, \quad
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \optr{L}^{p \alpha}_{q\beta} = E \sbra{q\beta}{\tilde E}\;.
\ee
In general, left and right eigenvectors do not coincide, but it is always possible to choose they normalization so that a orthonormality conditions among the
elements of left and right spaces holds, namely:
\be\lb{ExcitationOrthonormalityDef1}
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E}\sket{p\alpha}{E'} \equiv \dbraket{\tilde E}{E'} = \delta(E-E') \;.
\ee
Once that the eigenvectors are known, the spectral decomposition of the resolvent of $\optr{L}$ can be provided:  
\be
\left[\left(\bbomega-\optr{L}\right)^{-1}\right]^{p\alpha}_{q\beta} =\sint\dd E
\frac{\sket{p\alpha}{E}\sbra{q\beta}{\tilde E}}{\bbomega-E} \;,
\ee
so equation \eqref{LiouvilleTransBasisDef1} can be inverted:
\be\lb{RhopExcitationTransBasisDef1}
\sket{p\alpha}{\dm'(\omega)} = \sint\dd E \sket{p\alpha}{E}
\frac{\sum_q \sint \dd \beta\sbra{q\beta}{\tilde E}
\sket{q\beta}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;,
\ee
that can be compactly rewritten by using the index-less notation:
\be\lb{RhopExcitationTransBasisDef2}
\sket{p\alpha}{\dm'(\omega)} = 
\sint\dd E \sket{p\alpha}{E}
\frac{\dbraket{\tilde E}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;.
\ee

>>>>

The capability of the computational basis to express the results of the linear response can be analyzed using the localization argument previously discussed. 
The first step needed to address this problem concerns with the evaluation of the reliability of the $\mu$-basis for $\omega=0$, \emph{i.e.} for a perturbing field that does 
not depend on time. In this case the fluctuation states, as well as the expectation value of a given observable, can be computed performing a ground state calculation in which the perturbing 
field $\op\Phi$ is added to the external potential of $\hnot$\footnote{We are interested in the linear response, so the magnitude of $\op\Phi$ has to be chosen in a way that respects this 
regime. To check this point we can compute the perturbation-induced modification of an observable an verify that its value scales linearly with the magnitude of the field}. 
The goodness of the $\mu$-basis can thus be verified as in the standard ground state calculation, for instance by checking that, given a certain level of completeness of the basis,
the energy of the occupied (perturbed) orbitals is stable with respect to further refinements.

@@

We now introduce the notion of $\alpha$-basis, that is a set of orthonormal vectors $\{\ket{\psi(\alpha)}\}$ that form a basis in the subspace of the empty-states, \emph{i.e.} in the subspace 
orthogonal to the occupied eigenstates of $\hnot$. The elements of this set satisfy, by definition, the following orthonormality relation: 
\be\lb{psiAlphaDef1}
\brket{\psi(\alpha)}{\psi_p} = \brket{\psi_p}{\psi(\alpha)} = 0 \;, \qq
\brket{\psi(\alpha)}{\psi(\beta)} = \delta(\alpha-\beta) \;,
\ee
where we have parametrized the basis element through the multi-index $\alpha$, in the sense described in \ref{BasicDef}, since the associated vector can describe both a localized or a 
de-localized state. The $\op Q_0$ projector as well as the completeness relation can be easily written in function of this basis: 
\be\lb{psiAlphaQ0}
\op Q_0 = \sint\dd\alpha\ketbra{\psi(\alpha)}{\psi(\alpha)} \;, \;\; 
\identity = \sum_p\ketbra{\psi_p}{\psi_p} + \sint\dd\alpha\ketbra{\psi(\alpha)}{\psi(\alpha)} \;.
\ee
The $\alpha$-basis, defined through equations \eqref{psiAlphaDef1} and \eqref{psiAlphaQ0}, formally spans an infinite dimensional domain. However, for obvious computational reason it is 
interesting to express the $\{\ket{\psi(\alpha)}\}$ in the same $\mu$-basis used to perform the ground states calculations. The discretization conditions imposed by the $\mu$-basis reduce 
the $\{\ket{\psi(\alpha)}\}$ to a finite dimensional set. The dimension $n_\alpha$ of the basis can be computed by subtracting to the $n_\mu$ degrees of freedom the $n_{occ}$ constraints 
needed to imposes the orthonormality of the $\{\ket{\psi(\alpha)}\}$ to the occupied states, so that $n_\alpha=n_\mu-n_{occ}$.

\vspace{0.4cm}
\emph{Quality assessment of the reduced $\alpha$-basis}
\vspace{0.4cm}

When the sampling of the empty state subspace is needed to perform practical calculations in realistic systems the construction of a complete $\alpha$-basis typically turns out to be 
computationally unaffordable. In these cases it is possible to introduce a \emph{reduced} basis, denoted as $r\alpha$-basis, composed of a subset of dimension $n_{r\alpha}$ of the 
$\{\ket{\psi(\alpha)}\}$ elements of the complete basis.

The capability of the reduced basis to express the results of the linear response can be evaluated by using the results of a ground state calculation in which the time-independent 
perturbing field $\op\Phi$ is added to the external potential of $\hnot$. We denote with $\ket{\psi_p^\Phi}$ the occupied eigenstate of the perturbed Hamiltonian and, using the 
completeness relation of \eqref{psiAlphaQ0}, we can write the following expansion:
\be\lb{PertPsiExpansion1}
\ket{\psi_p^{\Phi}} = \sum_q a_{pq}\ket{\psi_q} + \sint\dd\alpha \,a_{p\alpha}\ket{\psi(\alpha)} \;,
\ee
where $a_{pq} = \brket{\psi_q}{\psi_p^{\Phi}}$ and $a_{p\alpha} = \brket{\psi(\alpha)}{\psi_p^{\Phi}}$. Equation \eqref{PertPsiExpansion1} holds exactly since the sum over the empty states
is extended to the complete $\alpha$-basis. Taking the norm of $\ket{\psi_p^\Phi}$ and summing over $p$ provides:
\be\lb{PertPsiExpansion2}
n_{occ} = \sum_{pq} |a_{pq}|^2 + \sum_p\sint\dd\alpha\,|a_{p\alpha}|^2 \;,
\ee
the two addend of this equation measure the distribution of the total electronic charge of the perturbed system between the occupied and empty subspaces. Furthermore, we can use
equation \eqref{PertPsiExpansion2} to build an estimator $C(n_{r\alpha})$ of the capability of the $r\alpha$-basis to describe the modification of the ground state density induced 
by the time-independent perturbation. We define this estimator as follows:
\be\lb{Cestimator1}
C(n_{r\alpha}) = 1 - \frac{\sum_p\sum_\alpha^{n_{r\alpha}}|a_{p\alpha}|^2}{n_{occ} - \sum_{pq} |a_{pq}|^2} \;,
\ee
$C(n_{r\alpha})$ is by construction a positive function with values between 1, achieved when the sampling of the empty-states subspace is absent and 0, obtained when the reduced basis
is complete. The normalization adopted in \eqref{Cestimator1} ensures that $C(n_{r\alpha})$ is an intensive estimator, so that its value does not depend on the number of occupied
orbitals of the system.

@@

The defining equation of the linear response can be reformulated using the explicit sampling of the empty states defined above. The fundamental observation on the basis
of this procedure regards the fact that both the response density operator \eqref{rhoPrimeOmegaDef1} and all the operators appearing in the Liouville equation 
\eqref{LiouvillianRhopomegaDef3} are expressed as a commutator with $\dmnot$. This observation implies that all the operators $\op O$ appearing in the contest of the linear 
response satisfy the formal relation:
\be
\op O = \dmnot\op O\op Q_0 + \op Q_0\op O\dmnot \;.
\ee
Plugging the operator expression of $\dmnot$ and $\op Q_0$ in terms of basis vector in the occupied and empty subspace provides:
\be\lb{OperatorTransBasisExpDef1}
\op O = \sum_p\sint \dd \alpha\left(
\excite{p}{\alpha}\trace{\decay{\alpha}{p}\op O} + \decay{\alpha}{p}\trace{\excite{p}{\alpha}\op O}
\right) \;,
\ee
where we have defined the \emph{transition operators}:
\be
\excite{p}{\alpha} \equiv \ketbra{\psi(\alpha)}{\psi_p}\;, \qq
\decay{\alpha}{p} \equiv \ketbra{\psi_p}{\psi(\alpha)} \;,
\ee
that thanks to \eqref{psiAlphaDef1} satisfy the orthonormality conditions:
\be\lb{transitionOpOrthonormalityDef1}
\trace{\excite{p}{\alpha}\excite{q}{\beta}} = \trace{\decay{\alpha}{p}\decay{\beta}{q}} = 0 \;, \qq
\trace{\excite{p}{\alpha}\decay{\beta}{q}} = \delta_{pq}\delta(\alpha-\beta) \;.
\ee
% \vspace{2cm}
% A generic operator can be decomposed in this basis and its formal expansion contains four distinct contribution of the form $\ketbra{\psi_p}{\psi_q}$, 
% $\ketbra{\psi(\alpha)}{\psi(\beta)}$, $\ketbra{\psi(\alpha)}{\psi_p}$, $\ketbra{\psi_p}{\psi(\alpha)}$. However an inspection of equations \eqref{rhoPrimeOmegaDef1}
% and \eqref{LiouvillianRhopomegaDef3} evidences that the relevant operators in the contest of the linear response are expressed as a commutator with $\dmnot$. 
% For this class of operators the only non vanishing contribution comes from the \emph{transition operators}:
% \be
% \excite{p}{\alpha} \equiv \ketbra{\psi(\alpha)}{\psi_p}\;, \qq
% \decay{\alpha}{p} \equiv \ketbra{\psi_p}{\psi(\alpha)} \;.
% \ee
% The orthonormality conditions \eqref{psiAlphaDef1} imply that:
% \be\lb{transitionOpOrthonormalityDef1}
% \trace{\excite{p}{\alpha}\excite{q}{\beta}} = \trace{\decay{\alpha}{p}\decay{\beta}{q}} = 0 \;, \qq
% \trace{\excite{p}{\alpha}\decay{\beta}{q}} = \delta_{pq}\delta(\alpha-\beta) \;,
% \ee
% so that the expansion of a suitable operator $\op O$ in the basis of transitions reads:
% \be\lb{OperatorTransBasisExpDef1}
% \op O = \sum_p\sint \dd \alpha\left(
% \excite{p}{\alpha}\trace{\decay{\alpha}{p}\op O} + \decay{\alpha}{p}\trace{\excite{p}{\alpha}\op O}
% \right) \;.
% \ee
Here we are assuming that the complete $\alpha$-basis is used, so the operator expansion in the basis of transition \eqref{OperatorTransBasisExpDef1} is exact. The effect of
the adoption of a reduced basis will be analyzed when needed. 
In order to recast this expansion in a simpler expression it turns out to be useful to define the rank-2 row vectors of the transition basis: 
\be
\sbra{p\alpha}{\cdot} \equiv 
\mat{ \excite{p}{\alpha} \;, & \decay{\alpha}{p} }\;,
\ee
and we also introduce a vectorial notation to deal with the coefficients of the expansion of an operator in the transition basis, that is: 
\be\lb{OperatorTransitionCoeffDef1}
\sket{p\alpha}{\op O}  \equiv \mat{ \trace{\decay{\alpha}{p}\op O } \\ 
\trace{\excite{p}{\alpha}\op O}} \; .
\ee
Thanks to this formalism we have that equation \eqref{OperatorTransBasisExpDef1} can be expressed as:
\be\lb{OperatorTransBasisExpDef2}
\op O = \sum_p\sint \dd \alpha \sbra{p\alpha}{\cdot}\sket{p\alpha}{\op O} = 
\dbraket{\cdot}{\op O} \;,
\ee
where the last equality implicitly define the ulterior compact notation of the ``big-shaped'' vectors in which the summed indexes are omitted.  
In the same fashion it will be also useful for the subsequent analysis to introduce a conjugate notation, in which the basis operators are expressed
as column vector and the coefficients of the expanded operator as row ones, that is: 
\be\lb{TransitionBasisConjugateDef1}
\sket{p\alpha}{\cdot} \equiv 
\mat{\decay{\alpha}{p} \\  \excite{p}{\alpha}} \;, \qq
\sbra{p\alpha}{\op O}  \equiv 
\mat{ \trace{\excite{\alpha}{p}\op O } \;, & \trace{\decay{p}{\alpha}\op O} }
\;,
\ee
where the flip in the role of $\op e$ and $\op d$ in \eqref{TransitionBasisConjugateDef1} follows from the properties of excitation and decay operators
under hermitian-conjugate operation. The expansion of an operator in the transition basis using the conjugate description reads: 
\be\lb{OperatorTransBasisConjugateExpDef1}
\op O = \sum_p\sint \dd \alpha \sbra{p\alpha}{\op O}\sket{p\alpha}{\cdot} = 
\dbraket{\op O}{\cdot} \;.
\ee

@@

%The Liouville equation \eqref{LiouvillianRhopomegaDef3} can be consistently written in the transition basis. 
% Indeed, its r.h.s is explicitly written
% as a commutator with $\dmnot$, while an inspection of the formal definition of the response density and of the Liouville superoperator evidences that
% they both contain a commutator with $\dmnot$. Written in the basis of transitions Liouville equation becomes a linear matrix equation for the rank-2 
% vectors of the coefficients \eqref{OperatorTransitionCoeffDef1} of the operators in its left and right side:
The Liouville equation \eqref{LiouvillianRhopomegaDef3} can be consistently written in the transition basis where it becomes a linear matrix equation for the 
rank-2 vectors of the coefficients \eqref{OperatorTransitionCoeffDef1} of the operators in its left and right side, that is:
\be
\omega\mat{ \trace{\decay{\alpha}{p}\dm'(\omega)} \\ \trace{\excite{p}{\alpha}\dm'(\omega)}} - 
\mat{ \trace{\decay{\alpha}{p}\Liouv\dm'(\omega)} \\ \trace{\excite{p}{\alpha}\Liouv\dm'(\omega)}} =
\mat{ \trace{\decay{\alpha}{p}\commutator{\op\Phi(\omega)}{\dmnot}} \\ \trace{\excite{p}{\alpha}\commutator{\op\Phi(\omega)}{\dmnot}}}
\;.\nn
\ee
The second addend in the l.h.s can be further analyzed by writing the transition basis expansion of $\dm'$ inside the traces. Performing this procedure
we can recast this quantity as the action of a 2-dimensional matrix on the rank-2 vector of the response density. So we have: 
\be\lb{LiouvilleTransBasisDef1}
\sum_q\sint\dd\beta \left[\omega\delta_q^p\delta(\alpha-\beta)- \optr{L}^{p \alpha}_{q\beta} \right]\sket{q\beta}{\dm'(\omega)} =
\sket{p\alpha}{\commutator{\op\Phi(\omega)}{\dmnot}} \;.
\ee
with:
\be\lb{LiouvilleMatrixTransBasisDef1}
\optr{L}^{p \alpha}_{q\beta} \equiv \mat{\trace{\decay{\alpha}{p}\, \Liouv\, \excite{q}{\beta}} &
\trace{\decay{\alpha}{p}\, \Liouv\, \decay{\beta}{q}} \\
\trace{\excite{p}{\alpha}\, \Liouv\, \excite{q}{\beta}} &
\trace{\excite{p}{\alpha}\, \Liouv\, \decay{\beta}{q}}} \;.
\ee

@@

This formulation of the Liouville equation is the starting point for the derivation of the spectral decomposition of the response density operator
in terms of the eigenvectors if the Liouville operator. 
The first step in the achievement of this result is provided by the construction of the eigenvectors of the matrix \eqref{LiouvilleMatrixTransBasisDef1}.
Following the notation previously defined we introduce the rank-2 vectors $\sket{p\alpha}{E}$ and $\sbra{p\alpha}{\tilde E}$ that represent, by definition,
the right and left eigenvectors $\optr{L}$, respectively, so that:
\be\lb{LiouvMatrixEigenvectorDef1}
\sum_q \sint \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q \beta}{E} = E \sket{p\alpha}{E} \;, \quad
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \optr{L}^{p \alpha}_{q\beta} = E \sbra{q\beta}{\tilde E}\;.
\ee
In general, left and right eigenvectors do not coincide, but it is always possible to choose they normalization so that a orthonormality conditions among the
elements of left and right spaces holds, namely:
\be\lb{ExcitationOrthonormalityDef1}
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E}\sket{p\alpha}{E'} \equiv \dbraket{\tilde E}{E'} = \delta(E-E') \;.
\ee
Once that the eigenvectors are known, the spectral decomposition of the resolvent of $\optr{L}$ can be provided:  
\be
\left[\left(\bbomega-\optr{L}\right)^{-1}\right]^{p\alpha}_{q\beta} =\sint\dd E
\frac{\sket{p\alpha}{E}\sbra{q\beta}{\tilde E}}{\bbomega-E} \;,
\ee
so equation \eqref{LiouvilleTransBasisDef1} can be inverted:
\be\lb{RhopExcitationTransBasisDef1}
\sket{p\alpha}{\dm'(\omega)} = \sint\dd E \sket{p\alpha}{E}
\frac{\sum_q \sint \dd \beta\sbra{q\beta}{\tilde E}
\sket{q\beta}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;,
\ee
that can be compactly rewritten by using the index-less notation:
% \be\lb{RhopExcitationTransBasisDef2}
% \sket{p\alpha}{\dm'(\omega)} = 
% \sint\dd E \sket{p\alpha}{E}
% \frac{\dbraket{\tilde E}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;.
% \ee
\be\lb{RhopExcitationTransBasisDef2}
\dket{\dm'(\omega)} = 
\sint\dd E \dket{E}
\frac{\dbraket{\tilde E}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;.
\ee
